---
title: "lab_report_knapsack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab_report_knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab6)
```

## package Intro:
The aim of this package is to study the effects of different algorithm with different computational complexity.

### 1.1 The knapsack 
The knapsack problem is a problem in combinatorial optimization: Given a set of items, each with a weight and a value, determine the number of each item to include in a collection so that the total weight is less than or equal to a given limit and the total value is as large as possible. It derives its name from the problem faced by someone who is constrained by a fixed-size knapsack and must fill it with the most valuable items.more details can be find in the link: [Knapsack Problem](https://en.wikipedia.org/wiki/Knapsack_problem) 

#### Solving:
there are several algorithms to solve, In this package Tree different algorithms with different computational complexity is defined and implemented.

* Algorithms:

   + Brute force search & Parallelize brute force search
   + Dynamic programming
   + Greedy heuristic
   
##### Brute force search & Parallelize brute force search
In this algorithm we are going through all possible candidates for the solution and checking whether each candidate satisfies the problem's statement.
The complexity of this algorithm is O(n^2).more details can be find in the link:
[Brute force search](https://en.wikipedia.org/wiki/Brute-force_search)

#### Parallelize code: 
When you have a list of repetitive tasks, you may be able to speed it up by adding more computing power. If each task is completely independent of the others, then it is a prime candidate for executing those tasks in parallel


#### Function Description:
* Function name: knapsack_brute_force(x, W)
* Input: x as a data.frame with two variable(v,w)
* Input: W Knapsack size
* Input: Parallel, if its FALSE, The Brute force search algorithm run and if Parallel is TRUE, The Parallelize brute force search run 
* Output:  returns the maximum knapsack value 
* Output: elements
note: In this function we use intToBits() to find all different combinations.

```{r, eval=TRUE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

library(parallel)
brute_force_knapsack<-function(x,W,parallel=FALSE){
  stopifnot(is.data.frame(x),W>0)
  n<-2^(length(x[,2]))-1
  d<-c()
  myfunc <- function(i){
    binary<-intToBits(i)
    d<-which(binary == 1)
    Len_d<-length(d)
    weight<-c()
    value<-c()
    for(j in 1:Len_d)
    {
      weight[j]<-x[,1][d[j]]
      value[j]<-x[,2][d[j]]
    }
    sum1<-sum(weight)
    if (sum1<=W)
    {
      return(sum(value))
    }else {
      return(0)
    }
    
  }
  if (parallel==FALSE){
    sum_value <- c(lapply(1:n, myfunc))
  }else{
    cores <- parallel::detectCores()
    cl <- parallel::makeCluster(cores, type = "PSOCK")
    sum_value <- parLapply(cl=cl, 1:n, fun = myfunc)
    # Shut down cluster
    parallel::stopCluster(cl)
  }
  

  max_sum_value <- sum_value[[which.max(sum_value)]]
  Sum_final<- round(max_sum_value)
  i1<-which( sum_value== max_sum_value)
  ele<-intToBits(i1)
  element<-which(ele == 1)
  Data<-list(value=Sum_final,elements=element)
  return(Data)
}

```
#### Question:
How much time does it takes to run the algorithm for n = 16 objects?

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

library(parallel)
brute_force_knapsack<-function(x,W,parallel=FALSE){
  stopifnot(is.data.frame(x),W>0)
  n<-2^(length(x[,2]))-1
  d<-c()
  myfunc <- function(i){
    binary<-intToBits(i)
    d<-which(binary == 1)
    Len_d<-length(d)
    weight<-c()
    value<-c()
    for(j in 1:Len_d)
    {
      weight[j]<-x[,1][d[j]]
      value[j]<-x[,2][d[j]]
    }
    sum1<-sum(weight)
    if (sum1<=W)
    {
      return(sum(value))
    }else {
      return(0)
    }
    
  }
  if (parallel==FALSE){
    sum_value <- c(lapply(1:n, myfunc))
  }else{
    cores <- parallel::detectCores()
    cl <- parallel::makeCluster(cores, type = "PSOCK")
    sum_value <- parLapply(cl=cl, 1:n, fun = myfunc)
    # Shut down cluster
    parallel::stopCluster(cl)
  }
  

  max_sum_value <- sum_value[[which.max(sum_value)]]
  Sum_final<- round(max_sum_value)
  i1<-which( sum_value== max_sum_value)
  ele<-intToBits(i1)
  element<-which(ele == 1)
  Data<-list(value=Sum_final,elements=element)
  return(Data)
}


system.time(brute_force_knapsack(x = knapsack_objects[1:16,], W = 3500))

```
#####  Brute force search & Parallelize brute force search
profiling Brute force search with parallelizing it, which can increase the speed of program running about 100%

```{r, eval=TRUE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

library(parallel)
brute_force_knapsack<-function(x,W,parallel=FALSE){
  stopifnot(is.data.frame(x),W>0)
  n<-2^(length(x[,2]))-1
  d<-c()
  myfunc <- function(i){
    binary<-intToBits(i)
    d<-which(binary == 1)
    Len_d<-length(d)
    weight<-c()
    value<-c()
    for(j in 1:Len_d)
    {
      weight[j]<-x[,1][d[j]]
      value[j]<-x[,2][d[j]]
    }
    sum1<-sum(weight)
    if (sum1<=W)
    {
      return(sum(value))
    }else {
      return(0)
    }
    
  }
  if (parallel==FALSE){
    sum_value <- c(lapply(1:n, myfunc))
  }else{
    cores <- parallel::detectCores()
    cl <- parallel::makeCluster(cores, type = "PSOCK")
    sum_value <- parLapply(cl=cl, 1:n, fun = myfunc)
    # Shut down cluster
    parallel::stopCluster(cl)
  }
  

  max_sum_value <- sum_value[[which.max(sum_value)]]
  Sum_final<- round(max_sum_value)
  i1<-which( sum_value== max_sum_value)
  ele<-intToBits(i1)
  element<-which(ele == 1)
  Data<-list(value=Sum_final,elements=element)
  return(Data)
}


```
#### Question:
How much time does it takes to run the algorithm for n = 16 objects?

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

library(parallel)
brute_force_knapsack<-function(x,W,parallel=FALSE){
  stopifnot(is.data.frame(x),W>0)
  n<-2^(length(x[,2]))-1
  d<-c()
  myfunc <- function(i){
    binary<-intToBits(i)
    d<-which(binary == 1)
    Len_d<-length(d)
    weight<-c()
    value<-c()
    for(j in 1:Len_d)
    {
      weight[j]<-x[,1][d[j]]
      value[j]<-x[,2][d[j]]
    }
    sum1<-sum(weight)
    if (sum1<=W)
    {
      return(sum(value))
    }else {
      return(0)
    }
    
  }
  if (parallel==FALSE){
    sum_value <- c(lapply(1:n, myfunc))
  }else{
    cores <- parallel::detectCores()
    cl <- parallel::makeCluster(cores, type = "PSOCK")
    sum_value <- parLapply(cl=cl, 1:n, fun = myfunc)
    # Shut down cluster
    parallel::stopCluster(cl)
  }
  

  max_sum_value <- sum_value[[which.max(sum_value)]]
  Sum_final<- round(max_sum_value)
  i1<-which( sum_value== max_sum_value)
  ele<-intToBits(i1)
  element<-which(ele == 1)
  Data<-list(value=Sum_final,elements=element)
  return(Data)
}


```

##### Dynamic programming
Dynamic Programming is mainly an optimization over plain recursion.
This method is based on simplifying a complicated problem by breaking it down into simpler sub-problems in a recursive manner.
The complexity of this algorithm is O(Wn).more details can be find in the link:
[Dynamic programming](https://en.wikipedia.org/wiki/Dynamic_programming)

#### Function Description:
* Function name: knapsack_dynamic(x, W)
* Input: x as a data.frame with two variable(v,w)
* Input: W Knapsack size
* Output:  returns the maximum knapsack value 
* Output: elements


```{r eval=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

knapsack_dynamic<-function(x, W){
  stopifnot(is.data.frame(x),W>0)
  weight_mat<-x[,1]
  Value_mat<-x[,2]
  n<-length(Value_mat)
  
  Dyn_mat<-matrix(0,nrow=n+1,ncol=W+1)
  for (i in 2:(n+1)){
    k<-(i-1)
    for(j in 2:(W+1)){
      q<-(j-1)
      if(weight_mat[k]>q){
        Dyn_mat[i,j]<-Dyn_mat[k,j]
      }else {
        M1<-Dyn_mat[k,j]
        M2<-Value_mat[k]
        M3<-(q-weight_mat[k])
        M4<-Dyn_mat[k,M3+1]
        M5<-M4+M2
        Dyn_mat[i,j]<-max(M1,M5)
      }
    }
    
  }
  Value_final<-Dyn_mat[i,j]
  Mat_final<-matrix(0,nrow=1,ncol=n)
  
  for (c in (n+1):2)
  {
    if(Dyn_mat[c,j]==Dyn_mat[c-1,j]){Mat_final[1,c-1]<-0
    } else {
      Mat_final[1,c-1]<-1
      j<-(W-weight_mat[c-1])+1
      
    }
  }
  elements<-which(Mat_final==1)
  value<-round(Value_final)
  Data<-list(value=value,elements=elements)
  return(Data)
}


knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
knapsack_dynamic(x = knapsack_objects[1:12,], W = 3500)
knapsack_dynamic(x = knapsack_objects[1:8,], W = 2000)
knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500))

```
#### Question:
How much time does it takes to run the algorithm for n = 500 objects?


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))
  

knapsack_dynamic<-function(x, W){
  stopifnot(is.data.frame(x),W>0)
  weight_mat<-x[,1]
  Value_mat<-x[,2]
  n<-length(Value_mat)
  
  Dyn_mat<-matrix(0,nrow=n+1,ncol=W+1)
  for (i in 2:(n+1)){
    k<-(i-1)
    for(j in 2:(W+1)){
      q<-(j-1)
      if(weight_mat[k]>q){
        Dyn_mat[i,j]<-Dyn_mat[k,j]
      }else {
        M1<-Dyn_mat[k,j]
        M2<-Value_mat[k]
        M3<-(q-weight_mat[k])
        M4<-Dyn_mat[k,M3+1]
        M5<-M4+M2
        Dyn_mat[i,j]<-max(M1,M5)
      }
    }
    
  }
  Value_final<-Dyn_mat[i,j]
  Mat_final<-matrix(0,nrow=1,ncol=n)
  
  for (c in (n+1):2)
  {
    if(Dyn_mat[c,j]==Dyn_mat[c-1,j]){Mat_final[1,c-1]<-0
    } else {
      Mat_final[1,c-1]<-1
      j<-(W-weight_mat[c-1])+1
      
    }
  }
  elements<-which(Mat_final==1)
  value<-round(Value_final)
  Data<-list(value=value,elements=elements)
  return(Data)
}

system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500))

```
`

##### Greedy heuristic
A greedy algorithm is any algorithm that follows the problem-solving heuristic of making the locally optimal choice at each stage.The complexity of this algorithm is O(nlog n).more details can be find in the link
[Greedy Algorithm](https://en.wikipedia.org/wiki/Greedy_algorithm)

#### Function Description:
* Function name: greedy_knapsack(x, W)
* Input: x as a data.frame with two variable(v,w)
* Input: W Knapsack size
* Output:  returns the maximum knapsack value 
* Output: elements
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))

greedy_knapsack<-function(x,W){
  stopifnot(is.data.frame(x),W>0)
  i<-1
  weight_mat<-x[,1]
  Value_mat<-x[,2]
  n<-length(Value_mat)
  mat_final_greedy<-c(Value_mat/weight_mat)
  mat_final_greedy1<-sort(mat_final_greedy,decreasing = TRUE)
  i<-1
  Value<-c()
  elements<-c()
  repeat  
  {
    d<-which(mat_final_greedy == mat_final_greedy1[i])
    W<- W-weight_mat[d]
    Value[i]<-Value_mat[d]
    elements[i]<-d
    i<-i+1
    if (W<0){
      break
    }
  }
  L<-length(Value)
  Value<-Value[1:L-1]
  Value<- round(sum(Value))
  L1<-length(elements)
  elements<-elements[1:L1-1]
  Data<- list(value=Value, elements=elements)
  return(Data)
}
greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
greedy_knapsack(x = knapsack_objects[1:1200,], W = 2000)
system.time(greedy_knapsack(x = knapsack_objects[1:1000000,], W = 2000))
```

#### Question:
How much time does it takes to run the algorithm for n = 1000000 objects?

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000))

greedy_knapsack<-function(x,W){
  stopifnot(is.data.frame(x),W>0)
  i<-1
  weight_mat<-x[,1]
  Value_mat<-x[,2]
  n<-length(Value_mat)
  mat_final_greedy<-c(Value_mat/weight_mat)
  mat_final_greedy1<-sort(mat_final_greedy,decreasing = TRUE)
  i<-1
  Value<-c()
  elements<-c()
  repeat  
  {
    d<-which(mat_final_greedy == mat_final_greedy1[i])
    W<- W-weight_mat[d]
    Value[i]<-Value_mat[d]
    elements[i]<-d
    i<-i+1
    if (W<0){
      break
    }
  }
  L<-length(Value)
  Value<-Value[1:L-1]
  Value<- round(sum(Value))
  L1<-length(elements)
  elements<-elements[1:L1-1]
  Data<- list(value=Value, elements=elements)
  return(Data)
}
system.time(greedy_knapsack(x = knapsack_objects[1:1000000,], W = 2000))
```


